Regression Cube Analysis of Cohort Study Data

### 2015-06-19 draft.tex
covered a bunch of reviewer comments

% -- My Own Packages and Commands
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\newcommand{\rem}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\add}[1]{\textcolor{blue}{#1}}
\newcommand{\com}[1]{\textcolor{red}{\underline{#1}}}
\newcommand{\design}[1]{\textcolor{orange}{#1}}

@@

Our contributions are:
\begin{itemize}
	%%
	\item An overview visualization technique representing feature interactions using target features.
	%%
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	%%
	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	%%
	\item The analysis of confounding features by providing comparison views between different formula results.
	%%
	\item The expansion and adaption of our method to any data with our open web-based system.
	%%
\end{itemize}
%%
The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
%%
\begin{enumerate}
	%%
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	%%
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
	%%
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
\design{
This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
}
\subsection{Iterative Design Based on Expert Feedback}
%%
\design{
%%
The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
%%
The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
%%
It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
%%
That is the focus of this work.
%%
% These target features often indicate the presence of the investigated disease.
%%
The domain experts wanted to \emph{model knowledge} about the investigated condition, such as confounding features (e.g., age or gender).
%%
For explorative analysis, they preferred an approach, which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
%%
Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
%%
This aspect is considered later in the Section~\ref{sec:SystemDesign}.
%%
%In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
}
%%
%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
%Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
%%
%Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
%In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
%These target features often indicate the presence of the investigated disease.
%%

@@

The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.

@@

By providing a service a website it has a much larger chance of being tested and potentially adapted by a wide range of experts.

@@

>>>>

% -- My Own Packages and Commands
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\newcommand{\rem}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\add}[1]{\textcolor{blue}{#1}}
\newcommand{\com}[1]{\textcolor{red}{\underline{#1}}}
\newcommand{\design}[1]{\textcolor{orange}{#1}}
\newcommand{\magenta}[1]{\textcolor{magenta}{#1}}
% white, black, red, green, blue, cyan, magenta, yellow.

@@

Our contributions are:

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand.
	%%
  \item An openly available web-based system, which incorporates a open data format to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
\noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
%%
\begin{enumerate}
	%%
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	%%
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
	%%
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
\magenta{
The \emph{Regression Cube} is designed to provide overview visualizations to support hypothesis generation.
%%
Hence, it is associated with step 1 and 2 of the epidemiological workflow presented in Section~\ref{sec:EpidemiologicalWorkflow}.
%%
Relationships observed using such techniques are subject of detailed statistical testing by statisticians with background in epidemiology using statistical processors, such as SPSS or STATA.
}
% \design{
% This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
% }
\subsection{Iterative Design Based on Expert Feedback}
%%
\design{
%%
The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
%%
The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
%%
It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
%%
That is the focus of this work.
%%
% These target features often indicate the presence of the investigated disease.
%%
The domain experts wanted to \emph{model knowledge} about the investigated condition, such as confounding features (e.g., age or gender).
%%
For explorative analysis, they preferred an approach, which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
%%
Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
%%
This aspect is considered later in the Section~\ref{sec:SystemDesign}.
%%
%In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
}
%%
%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
%Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
%%
%Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
%In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
%These target features often indicate the presence of the investigated disease.
%%

@@

The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses with other (statistical) tools.
\magenta{This is is, however, out of the scope of this work.}

@@

By providing a service a website it has a much larger chance of being tested and potentially adapted by a broad user base.

@@

\magenta{
%One limitation of the proposed method is that the regression metrics always only capture a part of the underlying model.
%%
%The analyst has to keep the aspects of the respective metric in mind to avoid false conclusions.
%%
The analysis is limited to three dynamic variables represented the cube dimensions.
%%
Investigating more dynamic variables can be achieved by projecting the resulting hyper-cube into a three-dimensional representation.
%%
This, however, increases the cognitive load and complexity of the analysis substantially and needs to be accompanied by techniques, which simplify this approach.
%%
Static features can be added at will using the formula input without increasing the complexity of the visualization.
}
